Background for transient faults in asynchronous circuits

### 2019-04-09 draft_together3.tex
finished current edit of proofs/ added in descriptions of each L and T in its beginning

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\section{Normal operation analysis}
As mentioned in the previous section on Speed Independent circuits, circuits can be modeled by a state graph.  
We want to define a trace, to describe the different paths a circuit can take as it follows the state graph.  Different paths can occur because there are states with multiple transitions from that state.  We record the succession of states and transitions by the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of infinite length where $s_i$ is the $i^{th}$ term of the sequence.  Formally this is written as $\sigma: \mathbb{N} \to S$ and $T(s_i,s_{i+1})$ for $\forall i \in \mathbb{N}$.  We also use the $\sigma_n$ as shorthand to denote the sequence of the first n states of $\sigma$$\end{definition} 
Even if there is only one transition from each state of a state graph, that is a linear state graph, our definition of trace is consistent with that.  \\
%add finite traces
% all states or just the state graph ones

[If the set of traces are equal then the state graph is equal]

Given two circuits, how can we determine when their basic functionality is the same even if they are implemented with different gates and circuit elements?  We define a notion of equivalence between two circuits.  We want the underlying state graph representation of the circuits to be equivalent.  However only the signals and the sequence of its transitions from the circuit is observable.  In other words the trace is observable.  Thus we define circuits as equivalent if the set of traces of the two circuits are the same. 
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B (up to a permutation of the variable names).  In other words, if it is possible to pair wires in A and B in such a way that $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}
%good enough for now, do I need if and only if? <-NO, xxx if xxx in math lingo is iff

%An additional property of the asynchronous circuit is that one can view each wire as a separate black box.  For a wire $v$, inputs from the rest of the circuit ($I_v$ and $I_v \subseteq$ V) feed into a gate and outputs the wire value $v$.  We call the instantaneous value of the gate for $v$ of a state $s$ as $f_v(s|_I)$, where f is the function $f:[I \to \{0,1\}] \to \{0,1\}$ and $s|_I$ is the state $s$ projected on the input variables $I_v$.  If inputs to the gate remain constant, the output $v$ will eventually take the value specified by $f_v(s|_I)$  %maybe change v again, also the f(s) vs f(I) thing...

@@

To prove claims about the fault tolerance of this circuit, we must first prove that the circuit behaves similar to the original circuit under normal operation.  That is, after applying a mapping of $h^A$, the set of traces of the resultant circuit and that of the original circuit are equal.  And likewise after applying a mapping of $h^B$.  We prove Lemma 1 which shows for each half of a circuit, if a transition occurs it must follow the transitions of the state graphs in the original circuit.  This property is independent of the signals on the other half circuit and is useful later in proving fault tolerance under a single transient fault.  %as long as it does not violate digital abstraction
However Lemma 1 does not guarantee that such a transition always exists so Lemma 2 is needed to show that a transition always exists and as a consequence the circuit cannot deadlock.  These two Lemmas combine together to show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\% and all traces in the original can be mapped to duplicated after h mapping?  <-this is obvious though by setting the A half and B half to both transition according to original trace

@@

One additional assumption that we make about the gates and wire values (including in the presence of faults) is that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the input wires.\\

@@

Note that the previous proof does not place any restrictions on the other half of the circuit other than that it does not break the digital abstraction for the C-elements that has inputs from both halves of the circuit (e.g. does not cause hazards in the C-element output).  Thus this Lemma is true regardless of whether there is a fault or not in the other half of the circuit.  It guarantees that if there is a next state it must follow the state transitions of the original circuit, but does not guarantee that there will be a next state.  %maybe move this to in front of L1
Next we will show that if both halves of the circuit start at the same valid initial state, deadlock does not occur.

@@

Combining the results of the previous Lemmas yield the following result:
\begin{theorem}[1]
For all traces $\sigma$ in the fault tolerant circuit, where $\sigma_n$ is the sequence of the first $n$ states of $\sigma$, then $h^{A,B}(\sigma_n)$ is a prefix of a trace in the original circuit for all $n \in \mathbb{N}$.  (h might make the trace shorter, so the prefix will not be of length n) %what was the point of the deadlock thing?
\end{theorem}

@@

@@

Next we will show that if a single transient fault occurs on any wire, the half circuit that does not contain the fault will still transition normally.  Applying Lemma 1, we know that the 'good' half will transition according to the original state graph but it may deadlock, so we must prove that it does not deadlock.  To do this we prove some properties that are true when a single transient fault occurs on a wire.  We partition the wires into two sets based on their properties when a single transient fault is added.  Lemma 3.1 shows that if a single fault occurs on the $c2$ or $c3$ wires for some wire $w$, the output wires still follow the original state graph.  Lemma 4 shows that if a fault occurs to any of the other wires, then the $c2$ or $c3$ wires follow a similar order to the non-fault case and thus the output wire pairs can never be different at deadlock.  Then we combine these lemmas to show that this implies that there is no deadlock.

@@

\begin{figure}
  \centering
    \includegraphics{flowl4c3v2}
  \caption{The wires in each row are in order of $c1$, $c2$ and $c3$.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  Note that case 5 transitions according to case 1 except with the 0's and 1's inverted.  The states transitions after that are inverted versions of state 2-4 until it transitions back to state 1 (uninverted).}
  \label{fig:l4}
\end{figure}

@@

In the induction step we assume $s_k$ is a state where each wire $w$ is in a case of figure \ref{fig:l4}, then at $s_{k+1}$ we look at the possible next states.  We don't assume anything about the non $c2$ and $c3$ wires for the A-half so that $c1^A$ may change values due to a fault.  However we do expect $c1^A$ to be either a 1 or 0 and that even with hazards on $c1^A$, $c2$ wires transition to either 0 or 1.  We indicate the possibly faulty $c1^A$ by the symbol X.
\begin{itemize}
	\item
Case 1:  $c2$, $c3$ are not excited.  $c1^B$ can be excited depending on the inputs to the logic block.  Therefore, if there is a transition from this state it must be to case 2.
\item
Case 2/3a/3b:  $c2$ can be excited depending on $c1^A$.  $c1^B$ cannot transition to 0 due to Semi-Modularity of wire $w$ and Lemma 1.  Therefore, if there is a transition it must be from case 2 to case 3a or 3b, and case 3a or 3b to case 4.
%I might need to show the excited ness of the good logic gate
\item
Cases 4/5a/5b:  $c1^B$ cannot transition to 0 due to the same reasoning above.  $c2$ is not excited. The only possible transitions are on the $c3$ wires and since there are no faults on them they must transition from case 4 to case 5a or 5b, and from case 5a or 5b to 6.
\end{itemize}
By induction, the property we set out to prove is true, the $c2$ and $c3$ wires will be in a case of figure \ref{fig:l4}. 
The same logic applies to when a fault occurs in the B half and the A half is fault free.  
\end{proof}
The importance of this property is that at most one of $c2$ or $c3$ pairs can be mismatched.  And thus all output pairs at deadlock must be equal.

\begin{theorem}[2]
When a single transient fault occurs, and for all $n \in \mathbb{N}$, if $\sigma$ is a trace in the fault tolerant circuit, then one of $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ (the non faulty half) is a prefix of a trace in the original circuit.  And the circuit does not deadlock.  
\end{theorem}
\begin{proof}
Without loss of generality we can assume the fault occurs in the A half of the circuit.  First by lemma 3 we know that if the fault occurs on $c2$ or $c3$ wire the circuit does not deadlock.  Thus we only look at when the fault occurs on all other wires. 
Next, we prove that a deadlocked circuit in this scheme must have at least one pair of output wires which have different values.  That is $w^A\neq w^B$ for some wire $w$.
This is shown by contradiction.  Assume that the circuit is deadlocked at state $s$ and $v^A = v^B$ for all wires $v$ from the original circuit.  If a transient occurs, one can wait until that wire is free to transition.  There
 must be a wire $w$ that is excited in the state $h^B(s)$ in the original circuit and we look at that wire block in the duplicated circuit.  We examine each case in lemma 4.
\begin{itemize}
	\item
	Case 1: Since $w$ is excited then by lemma 1 $c1^B$ or $w^B$ must be excited and can transition.
\item
Case 2:  $c1^A$ may be 0 or 1.  If $c1^A=0$, since the inputs to A half and B half are equal then $c1^A$ must be excited and can transition to 1.  If $c1^A=1$ then it can transition to case 3a or 3b.
\item
Case 3a/3b:  $c1^A$ may be 0 or 1.  If $c1^A=0$, since the inputs to A half and B half are equal then $c1^A$ must be excited and can transition to 1.  If $c1^A=1$ then it can transition to case 4.
\item
Case 4/5a/5b:  $c3$ is excited and can transition.
\end{itemize}
Thus a transition can always occur and the circuit is not deadlocked which leads to a contradiction.  To remain deadlocked, it must be true that there exists a wire where $w^A\neq w^B$, which implies its $c1^A\neq c1^B$ and $c2^A\neq c2^B$ and $c3^A\neq c3^B$.  But we have shown in Lemma 4 that $c2^A\neq c2^B$ and $c3^A\neq c3^B$ cannot occur at the same time under a single transient fault. Thus deadlock is not possible.  And applying Lemma 1, $h^B(\sigma_n)$ is a prefix of a trace in the original circuit. 
\end{proof}

>>>>

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\section{Traces and circuit equivalence}
We solve the problem of SEUs by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all its circuit elements.  Due to the asynchronous nature of how signals are processed, the resulting half circuits may have signals of corresponding wires at different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault tolerant circuit is equivalent to the original.  Finally we show that if an SEU occurs, one side of the fault tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\

Other than the basic terminology for asynchronous circuits and state graphs, we also want terminology to compare two circuits and especially evaluate whether they implement the same state graph.  
We first define a trace, to describe the different paths a circuit can take as it follows the state graph.  Different paths can occur because each state may have multiple transitions from that state.  We record the succession of states and transitions by the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of infinite length where $s_i$ is the $i^{th}$ term of the sequence.  Formally this is written as $\sigma: \mathbb{N} \to S$ and $T(s_i,s_{i+1})$ for $\forall i \in \mathbb{N}$.  We also use the $\sigma_n$ as shorthand to denote the sequence of the first n states of $\sigma$$\end{definition} 
Our definition of trace includes the case where there is only one transition from each state of a state graph, that is a linear state graph.  \\
%add finite traces
% all states or just the state graph ones

Given two circuits, how can we determine when their basic functionality is the same even if they are implemented with different gates and circuit elements?  We define a notion of equivalence between two circuits.  We want the underlying state graph representation of the circuits to be equivalent.  However only the signals and the sequence of its transitions from the circuit is observable.  In other words the trace is observable.  Thus we define circuits as equivalent if the set of traces of the two circuits are the same. [We use the assumption that if the set of traces are equal then the state graph is equal.]
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B (up to a permutation of the variable names).  In other words, if it is possible to pair wires in A and B in such a way that $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}
%good enough for now, do I need if and only if? <-NO, xxx if xxx in math lingo is iff

%An additional property of the asynchronous circuit is that one can view each wire as a separate black box.  For a wire $v$, inputs from the rest of the circuit ($I_v$ and $I_v \subseteq$ V) feed into a gate and outputs the wire value $v$.  We call the instantaneous value of the gate for $v$ of a state $s$ as $f_v(s|_I)$, where f is the function $f:[I \to \{0,1\}] \to \{0,1\}$ and $s|_I$ is the state $s$ projected on the input variables $I_v$.  If inputs to the gate remain constant, the output $v$ will eventually take the value specified by $f_v(s|_I)$  %maybe change v again, also the f(s) vs f(I) thing...
\section{Fault Tolerant transformation}

@@

\section{Normal operation analysis}
To prove claims about the fault tolerance of this circuit, we first prove that the circuit behaves similar to the original circuit under normal operation.  That is, after applying a mapping of $h^A$, the set of traces of the resultant circuit and that of the original circuit are equal.  And likewise after applying a mapping of $h^B$.  This goal is achieved by proving several Lemmas. Lemma 1 shows for each half of a circuit, if a transition occurs it must follow the transitions of the state graphs of the original circuit.  This property is independent of the signals on the other half circuit which is significant later in proving fault tolerance under a single transient fault as the half without the fault will transition as normal by Lemma 1.  %as long as it does not violate digital abstraction
One additional assumption that we make about the gates and wire values (including in the presence of faults) for Lemma 1 to hold is that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the input wires.\\

Note that Lemma 1 does not guarantee that a transition always exists from a given state so Lemma 2 is needed to show that a transition always exists and as a consequence the circuit cannot deadlock.  These two Lemmas combine together to show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\% and all traces in the original can be mapped to duplicated after h mapping?  <-this is obvious though by setting the A half and B half to both transition according to original trace

@@

This leads to Lemma 1, with the description and proof as follows:

@@

Note as described previously that the proof for Lemma 1 does not place any restrictions on the other half of the circuit other than that it does not break the digital abstraction for the C-elements that has inputs from both halves of the circuit (e.g. does not cause hazards in the C-element output).  Thus this Lemma is true regardless of whether there is a fault or not in the other half of the circuit.  It guarantees that if there is a next state it must follow the state transitions of the original circuit, but does not guarantee that there will be a next state.\\  %maybe move this to in front of L1

This leads to the next property that we will show: if both halves of the circuit start at the same valid initial state, the wire blocks of the circuit transitions in a completely predictable pattern and thus deadlock does not occur.

@@

Combining the results of the previous Lemmas shows that the fault tolerant circuit is equivalent to the original circuit under normal operation.  This is defined formally in a theorem as follows:
\begin{theorem}[1]
For all traces $\sigma$ in the fault tolerant circuit, $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ is a prefix of a trace in the original circuit for all $n \in \mathbb{N}$.  (h might make the trace shorter, so the prefix will not be of length n) %what was the point of the deadlock thing?
\end{theorem}
Apply Lemma 1 then $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ is a prefix of a trace.  From Lemma 2.2, no deadlock occurs so it is true for all $n\in \mathbb{N}$.

@@

Next we consider what happens to the fault tolerant circuit under a single transient fault which is defined as follows:

@@

Next we will show that if a single transient fault occurs on any wire, the fault free half circuit will be equivalent to the original circuit.  Applying Lemma 1, we know that the fault free half will transition according to the original state graph but it may deadlock, so we must prove that it does not deadlock.  To do this we prove some properties that are true when a single transient fault occurs on a wire.  We partition the wires into two sets based on their properties when a single transient fault is added.  Lemma 3 shows that if a single fault occurs on the $c2$ or $c3$ wires for some wire $w$, the output wires still follow the original state graph.  Lemma 4 shows that if a fault occurs to any of the other wires, then the $c2$ or $c3$ wires follow a similar order to the non-fault case and thus the output wire pairs can never be different at deadlock.  Then we combine these Lemmas to show that there is no deadlock.

@@

\begin{figure}
  \centering
    \includegraphics{flowl4c3v2}
  \caption{The wires in each row are in order of $c1$, $c2$ and $c3$.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  Note that case 5 transitions according to case 1 except with the 0's and 1's inverted.  The states transitions after that are inverted versions of state 2-4 until it transitions back to state 1 (uninverted).}
  \label{fig:l4}
\end{figure}
Next we show that if a fault occurs to any of the other wires, then the $c2$ or $c3$ wires of each wire block of the circuit transitions in a completely predictable pattern analogous to Lemma 2.1.  The main difference is that because a fault may be present we cannot fully predict the value of one of $c1^A$ or $c1^B$.  The following property is true despite not knowing one of the $c1$ wire values and leads to the conclusion that the output wire pairs can never be different at deadlock.

@@

In the induction step we assume $s_k$ is a state where each wire $w$ is in a case of figure \ref{fig:l4}, then at $s_{k+1}$ we look at the possible next states.  We don't assume anything about the non $c2$ and $c3$ wires for the A-half so that $c1^A$ may change values due to a fault.  However we do expect $c1^A$ to be either a 1 or 0 and that even with hazards on $c1^A$, $c2$ wires transition to either 0 or 1.  We indicate the possibly faulty $c1^A$ by the symbol X.
\begin{itemize}
	\item
Case 1:  $c2$, $c3$ are not excited.  $c1^B$ can be excited depending on the inputs to the logic block.  Therefore, if there is a transition from this state it must be to case 2.
\item
Case 2/3a/3b:  $c2$ can be excited depending on $c1^A$.  $c1^B$ cannot transition to 0 due to Semi-Modularity of wire $w$ and Lemma 1.  Therefore, if there is a transition it must be from case 2 to case 3a or 3b, and case 3a or 3b to case 4.
%I might need to show the excited ness of the good logic gate
\item
Cases 4/5a/5b:  $c1^B$ cannot transition to 0 due to the same reasoning above.  $c2$ is not excited. The only possible transitions are on the $c3$ wires and since there are no faults on them they must transition from case 4 to case 5a or 5b, and from case 5a or 5b to 6.
\end{itemize}
By induction, the property we set out to prove is true, the $c2$ and $c3$ wires will be in a case of figure \ref{fig:l4}. 
The same logic applies to when a fault occurs in the B half and the A half is fault free.  
\end{proof}
The importance of this Lemma is that at most one of $c2$ or $c3$ pairs can be mismatched.  The next theorem shows that this implies all output pairs at deadlock must be equal so there is a possible next condition which contradicts the definition of deadlock.  Thus the fault tolerant circuit does not deadlock under a single transient fault.

%Then the following corollary holds:
%\begin{corollary}[4.2]
%All output pairs at deadlock must be equal
%\end{corollary}

\begin{theorem}[2]
When a single transient fault occurs, if $\sigma$ is a trace in the fault tolerant circuit, then one of $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ (the non faulty half) for all $n \in \mathbb{N}$ is a prefix of a trace in the original circuit.  And the circuit does not deadlock.  
\end{theorem}
\begin{proof}
Without loss of generality we can assume the fault occurs in the A half of the circuit.  First by Lemma 3 we know that if the fault occurs on $c2$ or $c3$ wire the circuit does not deadlock.  Thus we only look at when the fault occurs on all other wires. 
Next, we prove that a deadlocked circuit in this scheme must have at least one pair of output wires which have different values.  That is $w^A\neq w^B$ for some wire $w$.
This is shown by contradiction.  Assume that the circuit is deadlocked at state $s$ and $v^A = v^B$ for all wires $v$ from the original circuit.  If a transient occurs, one can wait until that wire is free to transition.  There
 must be a wire $w$ that is excited in the state $h^B(s)$ in the original circuit and we look at that wire block in the duplicated circuit.  We examine each case in Lemma 4.
\begin{itemize}
	\item
	Case 1: Since $w$ is excited then by Lemma 1 $c1^B$ or $w^B$ must be excited and can transition.
\item
Case 2:  $c1^A$ may be 0 or 1.  If $c1^A=0$, since the inputs to A half and B half are equal then $c1^A$ must be excited and can transition to 1.  If $c1^A=1$ then it can transition to case 3a or 3b.
\item
Case 3a/3b:  $c1^A$ may be 0 or 1.  If $c1^A=0$, since the inputs to A half and B half are equal then $c1^A$ must be excited and can transition to 1.  If $c1^A=1$ then it can transition to case 4.
\item
Case 4/5a/5b:  $c3$ is excited and can transition.
\end{itemize}
Thus a transition can always occur and the circuit is not deadlocked which leads to a contradiction.  To remain deadlocked, it must be true that there exists a wire where $w^A\neq w^B$, which implies its $c1^A\neq c1^B$ and $c2^A\neq c2^B$ and $c3^A\neq c3^B$.  But we have shown in Lemma 4 that $c2^A\neq c2^B$ and $c3^A\neq c3^B$ cannot occur at the same time under a single transient fault. Thus deadlock is not possible.  And applying Lemma 1, $h^B(\sigma_n)$ is a prefix of a trace in the original circuit. 
\end{proof}
